Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-11 paper/disc20clustering.tex
wip

In this section, we consider the learning variant of dynamic balanced graph partitioning problem.
For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
The lower bound holds also in the general partitioning model (studied in Section~\ref{sec:part}).
At the end of this section, we discuss an asymptotically optimal upper bound for the learning variant.

@@

We provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
Later we elaborate on how to efficiently transform it to a lower bound for the general partitioning problem.
The lower bound requires $k\geq 3$.
For $k=2$, the learning problem is trivial: immediate collocation of a communicating pair is $1$-competitive.
(A partitioning problem for $k=2$ is non-trivial, a lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)

@@

A~\emph{ground set} is a group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
Their role is to ensure that the algorithm maintains a~partition of ground sets into clusters.
We emphasize that ground sets are revealed only upon their nodes are split.
The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
Once we construct a ground set, it lasts until the end of the input sequence.
We say that a ground set is \emph{singleton} if it contains exactly one node.
We say that the node that belongs to a singleton ground set is \emph{isolated}.

>>>>

In this section, we consider the learning variant of dynamic balanced graph partitioning problem.
For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$ for $k\geq 3$.
The lower bound holds also in the general partitioning model (studied in Section~\ref{sec:part}).
At the end of this section, we discuss an asymptotically optimal upper bound for the learning variant.

@@

We provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
Later we elaborate on how to efficiently transform it to a lower bound for the general partitioning problem.
The lower bound requires $k\geq 3$.
In contrast, for $k=2$ the learning problem is trivial: immediate collocation of communicating pairs is $1$-competitive.
(A partitioning problem for $k=2$ is non-trivial, a lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)

@@

A~\emph{ground set} is a group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
Their role is to ensure that the algorithm maintains a~partition of ground sets into clusters.
We emphasize that ground sets are revealed only upon their nodes are split.
The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
Once we construct a ground set, it lasts until the end of the input sequence.
We say that a ground set is a \emph{singleton} if it contains exactly one node.
We say that the node that belongs to a singleton ground set is \emph{isolated}.
